---
title: "Enter your title here"
author: "Enter your name here"
date: "11/15/2020"
output: 
  html_document:
    toc: true
    number_section: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Introduce your project. You can use what you submitted in your project proposal. Organize it in paragraphs (do not include guiding questions).

# Research Questions

Clearly state your motivation for your research questions (why you are asking what you are asking) and state your research question.

1. First research question
1. Second research question

# Methods

This study seeks to answer the research questions above through a quantitative data analysis. In this section, I describe the data used in this study, and the variables used to answer my research question.

## Data Description

```{r}
# read in data
# glimpse or other data summary
```

# Results

Include here your tables, plots, and prose that explains what we can learn from these tables and plots. Discuss how your tables and plots answer your research question.

# Conclusion

Re-state your research question, and how your analysis results answer the research question. Mention any limitations you can think of. Close the conclusion with futher directions and/or next steps.

# References

Add here any cited works
